---
title: "US Lending Club 2018 Q1 to 2019 Q2"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    social: menu
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(data.table)
library(tidyverse)
library(lubridate)
library(plotly)
library(crosstalk)
library(shiny)
library(dbplyr)
library(vcd)
options(shiny.sanitize.errors = FALSE)

setwd("D:/Data Science/CETM 25 Data Visualisation/Assignment")
y2018<-c("2018 Q1", "2018 Q2", "2018 Q3", "2018 Q4")
loan_raw <- fread("loan.csv")
loan2018_raw <- loan_raw[year_quarter%chin%y2018]
loan2019_raw <- loan_raw[year_quarter%chin%c("2019 Q1", "2019 Q2"),]
state_to_abbr<- fread("state.csv")
loan2018<-loan2018_raw
loan2019<-loan2019_raw
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

```

Introduction
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------

### Loan by States

```{r}
loan_map <- loan_raw[loan_status %chin% c("Charged Off", "Current", "Issued", "In Grace Period", 
                                              "Late (16-30 days)", "Late (31-120 days)"), 
                                         .(no_app = .N, total_int=sum(total_int), 
                                           mn_an_income=round(mean(annual_inc), 2), 
                                          loan_amount = sum (loan_amnt), 
                                          mn_int_rate=round(mean(loan_int),4)), by="addr_state"]
loan_map <- left_join(state_to_abbr, loan_map, by = c("code" = "addr_state"))
loan_map$hover <- with(loan_map, paste(state, '<br>', 
                                       "No of App",format(no_app, big.mark = ","), '<br>', 
                                       "Issued loan Amount", format(loan_amount, big.mark = ","),'<br>',
                                       "Expected Revenue", format(total_int, big.mark = ","), '<br>', 
                                       "Average Annual Income", format(mn_an_income, big.mark = ","), '<br>', 
                                       "Average Interest Rate", mn_int_rate*100,"%"))

map_plot <- plot_geo(loan_map, locationmode = 'USA-states') %>%
  add_trace(
    z = ~total_int, text = ~hover, locations = ~code,
    color = ~total_int, colors = 'Blues'
  ) %>%
  colorbar(title = "Expected Revenue in USD") %>%
  layout(
    title = 'Expected Revenue (Interests) on Each State',
    geo = g
  )

map_plot
```

### Bad Loan

```{r}
default_map <- loan_raw[bad_loan %chin% "Bad Loan", .(no_app = .N, total_int=sum(total_int), mn_an_income=round(mean(annual_inc), 2), 
                                                 loan_amount = sum (loan_amnt), mn_dti=round(mean(dti, na.rm=TRUE),4)), by="addr_state"]
default_map <- left_join(state_to_abbr, default_map, by = c("code" = "addr_state"))
default_map$hover <- with(default_map, paste(state, '<br>', 
                                             "No of Bad Loans: ",format(no_app, big.mark = ","), '<br>', 
                                             "Total Loan Interest Involoved: ", format(total_int, big.mark = ","),'<br>',
                                             "Total Bad Loan Amount: ", format(loan_amount, big.mark = ","),'<br>',
                                             "Average Annual Income", format(mn_an_income, big.mark = ","), '<br>', 
                                             "Average Debt-to-Income Ratio", mn_dti*100))

default_map_plot <- plot_geo(default_map, locationmode = 'USA-states') %>%
  add_trace(
    z = ~total_int, text = ~hover, locations = ~code,
    color = ~total_int, colors = 'Reds'
  ) %>%
  colorbar(title = "Bad Loan Interest in USD") %>%
  layout(
    title = 'Bad Loan on Each State',
    geo = g
  )

default_map_plot
```

Column {data-width=250}
-----------------------------------------------------------------------

### Year-Quarter Table

```{r}
table_loan <- plot_ly(
  type = 'table',
  header = list(
    values = c('<b>Year-Quarter</b>', '<b>2018 Q1-Q4</b>','<b>2019 Q1-Q2</b>'),
    line = list(color = '#506784'),
    fill = list(color = '#119DFF'),
    align = c('center','center'),
    font = list(color = 'white', size = 12)
  ),
  cells = list(
    values = rbind(
      c('No. of Applications', 'Expected Revenue (USD)', 'Average Annual Income (USD)', 'Total Loan Amount (USD)', 'Average Interest Rate', 'No. of Bad Loans', 'Default Rate', 'Bad Loan Amount (USD)'),
      c("442,629","2,031M","80,078.42","7,097M","12.72%","27,706", "0.16%", "482.77M"),#2018
      c("246,814","1,165M","80,046.67","4,103M","12.73%","3,166", "0.02%", "51.14M")),#2019
    line = list(color = '#506784'),
    fill = list(color = c( 'white')),
    align = c('center', 'center'),
    font = list(color = c('#506784'), size = 12)
  ))
table_loan

```


### Achievement of Issued Loan Amount in 2019

```{r}
rate <- round(sum(loan2019$loan_amnt)/sum(loan2018$loan_amnt)*100,2)
gauge(rate, min = 0, max = 100, symbol = '%', gaugeSectors(
  success = c(80, 100), warning = c(40, 79), danger = c(0, 39)))
```




Year-Quarter {data-navmenu="Loan by Types"}
=======================================================================

Column
-----------------------------------------------------------------------

### Year-Quarter

```{r}
quarter_loan<-loan_raw[bad_loan %chin% "Bad Loan", .(aver_int=round(mean(loan_int)*100, 2), aver_loan_amnt=mean(loan_amnt)), by=c("year_quarter", "loan_status")]

quarter_plot<- quarter_loan%>%
                  plot_ly(x=~year_quarter,y=~aver_loan_amnt,
                          hoverinfo="text", 
                          text=~paste("Average Interest Rate: ", round(aver_int, 2), "<br>",
                                      "Average Loan Amount: ", round(aver_loan_amnt, 2)))%>%
                          add_lines(linetype=~loan_status,color = ~loan_status)%>%
                          layout(xaxis=list(title="Year Quarter"),
                                 yaxis=list(title="Average Loan Amount"),
                                 title="Average Loan Amount on Loan Purpose by Income Category")  
            
quarter_plot
```

Brief Study on Bad Loans {data-navmenu="Loan by Types"}
=======================================================================

Column {.tabset}
-----------------------------------------------------------------------

### Credit Grades

```{r}
grade_loan_pie<-loan_raw[bad_loan%chin%"Bad Loan",.(no_of_loan=.N), by="grade"]
grade_plot_pie<-grade_loan_pie%>%
  plot_ly(labels=~grade, values=~no_of_loan,type='pie',
                    #testposition='inside',
                    textinfo = 'label+percent',
                    insidetextfont = list(color = '#FFFFFF'),
                    hoverinfo = 'text',
                    text = ~paste('Number of Loans: ', no_of_loan),
                    marker = list(colors = ~grade,
                                  line = list(color = '#FFFFFF', width = 1)),
                    #The 'pull' attribute can also be used to create space between the sectors
                    showlegend = FALSE) %>%
                        layout(title = 'Credit Grade Proportion on Bad Loan',
                    xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                    yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
grade_plot_pie

```

### Income Categories

```{r}
income_cat_loan_pie<-loan_raw[bad_loan%chin%"Bad Loan",.(no_of_loan=.N), by="income_cat"]
income_cat_plot_pie<-income_cat_loan_pie%>%
  plot_ly(labels=~income_cat, values=~no_of_loan,type='pie',
                    #testposition='inside',
                    textinfo = 'label+percent',
                    insidetextfont = list(color = '#FFFFFF'),
                    hoverinfo = 'text',
                    text = ~paste('Number of Loans: ', no_of_loan),
                    marker = list(colors = ~income_cat,
                                  line = list(color = '#FFFFFF', width = 1)),
                    #The 'pull' attribute can also be used to create space between the sectors
                    showlegend = FALSE) %>%
                        layout(title = 'Income Categories Proportion on Bad Loa',
                    xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                    yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
income_cat_plot_pie
```

### Regions

```{r}
region_loan_pie<-loan_raw[bad_loan%chin%"Bad Loan",.(no_of_loan=.N), by="region"]
region_plot_pie<-region_loan_pie%>%
  plot_ly(labels=~region, values=~no_of_loan,type='pie',
                    #testposition='inside',
                    textinfo = 'label+percent',
                    insidetextfont = list(color = '#FFFFFF'),
                    hoverinfo = 'text',
                    text = ~paste('Number of Loans: ', no_of_loan),
                    marker = list(colors = ~region,
                                  line = list(color = '#FFFFFF', width = 1)),
                    #The 'pull' attribute can also be used to create space between the sectors
                    showlegend = FALSE) %>%
                        layout(title = 'Region Proportion on Bad Loa',
                    xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                    yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
region_plot_pie

```

Loan Purpose {data-navmenu="Loan by Types"}
=======================================================================

Column
-----------------------------------------------------------------------

### Loan Purposes

```{r}
title_plot<- loan_raw%>%
                group_by(title, income_cat, year_quarter)%>%
                summarise(aver_int=mean(loan_int)*100, aver_loan_amnt=mean(loan_amnt))%>%
                plot_ly(x=~aver_int,y=~title,
                        hoverinfo="text", 
                        text=~paste("Average Interest Rate: ", round(aver_int, 2), "<br>",
                                    "Average Loan Amount: ", round(aver_loan_amnt, 2)))%>%
                  add_text(
                  x=22,y="Other",text=~year_quarter, frame=~year_quarter,
                  textfont=list(size=75, color=toRGB("gray80")))%>%
                  add_markers(frame=~year_quarter,   
                              color=~income_cat,colors=c("#0040FF", "#FE9A2E", "#FE2E2E"),
                              size=~aver_loan_amnt, marker=list(sizemode="diameter"))%>%
                  layout(xaxis=list(title="Average Interest Rate"),
                         yaxis=list(title="Loan Purpose"),
                         title="Average Interest Rate on Loan Purpose by Income Category")%>%
                  animation_slider(hide=TRUE)
title_plot               
                 
```

Region {data-navmenu="Loan by Types"}
=======================================================================

Column {.sidebar}
-----------------------------------------------------------------------

```{r, }
title_type<- c("North West", "South East", "West", "Mid West", "South West")
checkboxGroupInput(
        "region",
        "Regions:",
        choices = title_type,
        selected = title_type
      )

```


Column
-----------------------------------------------------------------------

### By Regions

```{r}
region_loan <- loan_raw%>%
                  group_by(home_ownership, region)%>%
                  summarise(avg_annual_inc = round(mean(annual_inc),2))
  reactive_region <- reactive(
    region_loan %>%
      filter(region %in% input$region)
  )

renderPlotly({
    validate(
      need(input$region, 'Check at least one region!')
    )
    reactive_region() %>% 
      plot_ly(x=~home_ownership, y=~avg_annual_inc, color=~region)%>%
      add_bars() %>%
      layout(xaxis=list(title="Home Ownership"), 
             yaxis=list(title="Average Annual Income"))
  })
```


Assoication
=======================================================================

Column {.sidebar}
-----------------------------------------------------------------------

>This moasic plot indicates the association between credit grades and class of income.

>For each loan purpose selection, the plot would show the specific findings on selected loan purpose. 

>Signs at the right reveal the Peason's residuals. Blue means positively assoicated; Red means negatively associated.



```{r}
vcd_loan <- loan_raw[,c("grade", "title", "income_cat")]
title_type<- c("Business", "Car financing", "Credit card refinancing", "Debt consolidation", 
               "Green loan", "Home buying", "Home improvement", "Major purchase", "Medical expenses",
               "Moving and relocation", "Other")
selectInput("title","Loan Purpose:", choices = title_type, selected = "Business")

```




Column
-----------------------------------------------------------------------

### Mosaic Plot
```{r}
renderPlot({
  vcd_loan<-vcd_loan[title%chin%input$title]
  mosaic(~grade+income_cat, data = vcd_loan,
         shade = TRUE, split_vertical = c (TRUE, FALSE))
  
})
```

